Supervised Learning for Psychological Attention to High School Students

### 2020-11-18 article/article.tex
[ article ] Formatted and corrected typos.

Given that the sample size is fixed, the analysis for the ``Probably
Approximately Correct'' (PAC) learning was not straight-forward, as the
classifier's maximum permited error $\epsilon$ was estimated from the classic
formulae, rather than calculating the adequate sample size for a given $\delta$.

The PAC learning here applied is based on the ideas presented in
\parencite{kubat2017}. Let $N$ be number of samples in the training set,
$\epsilon$ is the classifier maximum permited error rate, $\delta$ is the
probability that a classifier with error rate greater than $\epsilon$ is
\text{error-free} on the training set, $m$ be the dimension of each training
patttern, and let $VC$ be the Vapnik-Chervonenkis dimension of the given
classifier. Hence, for learning in continuous domains, the lower bound for
$N$ is

@@

Furthermore, it is important to notice that the dataset is 6-dimensional, which
clearly brings problems for visualizing the learning machines' outputs and
decision boundaries. In this paper, the decision functions of the fitted
learning machines will be presented using contour plots of the projected data.
This process will better described: each 6-dimensional pattern generates an
output whenever it is evaluated in the fitted model. The idea is to generate an
equally-spaced grid in $[-1,1]^6$ and evaluate each point on this model, in
order to plot the contour of the shattering function. The main issue is that a
direct projection into a 2-dimensional space of each point on the grid is not
necessarily a function, since several points can have the same projection, but
different values of the shattering function. In order to address this issue,
the mean of the points that mapped into the same projection is taken, giving an
proxy of the general behavior in high dimensions. This ``mean'' projection
contour plots also include the projection of the training data points into the
same coordinates and scattered with the real labels, giving an idea of the
classification performance in high dimensions. Remark: as stated, the survey
data also includes the outputs of the Fuzzy Inference System; thus, the countour
plots are presented only for relating first input with the first input, second
with the second output and so on. This was done aiming at two objectives:
keeping a reasonable number of plots in the document, and giving a more
intuitive interpretation to these plots.

@@

The same idea from the SVM classifiers for plotting the countour of the decision
boundaries of the fitted model was used, plotting the training dataset with the
real labels and the contour for the decision function.

Additionally, it is well-known that the bound for the training dataset size is
given by
\begin{equation}
  N\geq \dfrac{\ln(2)}{2\epsilon^2}\left[(2^k-1)(1+\log_2 m) + 1 + \ln\left(
  \dfrac{1}{\delta}\right)\right],
\end{equation}
but again, the size of the dataset is fixed. Thus, $\epsilon$ was calculated by
fixing the tree depth $k=4$ and setting $\delta=0.3$ yielding a value of
$\epsilon=0.5381$, which is a suitable value for the classifier error.

@@

The results are presented using three learning curves: the sum of local
gradients per layer at the end of each epoch, the average error energy and a
Receiver Operating Characteristic (ROC) curve, comparing the obtained results
with the real labels.
\begin{itemize}
  \item The sum of gradients per layer at the end of each epoch is an approach
  visualizing if the training algorithm is solving the optimization problem,
  since the sum of local gradients must converge to zero as each gradient
  converges to zero. It is well-known that the vanish of gradients is a
  necessary condition for optimality and the training algorithm is just an
  optimization problem minimizing the output error (cost function).
  \item The average error energy is the cost function of the optimization
  problem. Let $y_i\in\mathbb{R}^n$ be the real output of a given input pattern
  $x_i\in\mathbb{R}^m$ and let $\hat{y}_i\in\mathbb{R}^n$ be the output of the
  NN for the same input pattern. Denoting each component by superscripts, the
  average error energy $\xi_{av}$ on each epoch is then computed as
\begin{equation}
\xi_{av}=\dfrac{1}{2}\sum_{i=1}^{N}\sum_{k=1}^n\left(y^k_i-\hat{y}^k_i\right)^2,
\end{equation}
  where $N$ is the training sample size. Thus, it is clear that
  $\xi_{av}\rightarrow0$ is a good initial approach for evaluating the
  learning performance.
  \item The ROC curve is a tool that helps visualize the correct classifications
  of the learning machine by plotting the true positive rate on the vertical
  axis, and false positive rate on the horizontal axis. The Area Under the Curve
  (AUC) is an indicator of how good the data is being classified by the model,
  with higher values being usually better.
\end{itemize}

\subsection{Sensitivity-Specificity}
The sensitivity is a ``measure of the proportion of actual positive cases that
got predicted as postive (true positive)'' and the specificity is ``measure of
proportion of actual negatives that got predicted as negatives (true negatives)
'' \parencite{sens}. This values are, hence, calculated using the formulae
\begin{align}
  \mathrm{Sensitivity}=&\dfrac{TP}{TP+FN},\quad
  \\\mathrm{Specificity} =&
  \dfrac{TN}{TN+FP},
\end{align}
where TP stands for True Postives, FN for False Negatives, TN for True Negatives
and FP for False Negatives, which are extracted from a confusion matrix.

\section{Results}\label{sec:res}
For giving an idea of the dataset treated in this work (visualization), the data
was embedded into a 2-dimensional space using the T-student Stochastic
Neighborhood Embedding (TSNE) \href{https://github.com/lejon/TSne.jl}{in Julia}
\parencite{maaten2008}. In Fig. \ref{fig:emb_dat}, the embedded dataset is
presented.
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/embedded-data.pdf}
  \caption{Embedded dataset.}
  \label{fig:emb_dat}
\end{figure}

@@

\subsection{Neural Networks}
Recall that only the worst and 2 best NN are kept for presenting the results.
The worst NN (in the sense of the objective function) was trained with
$\eta=0.9$ and it has a $\mathcal{L}=[1,2,3]$ architecture in the hidden layers.
The learning curves for this MLP are presented in Fig. \ref{fig:NN-worst}.
This architecture is expected to perform poorly, at least for the given number
of epochs, since it is attempting to summarize all 6 inputs in a one dimensional
value for the first hidden layer and then casting this value into higher
dimensions. It is unkown for the authors if this machine can learn the data at
all, since the behavior of the backpropagation algorithm is unpredictable as
epochs advance. Note from Fig. \ref{subfig:grad-worst-NN}, the gradients are
stable in the epoch horizon defined, from Fig. \ref{subfig:error-worst-NN} that
the average energy error seems to reach a stable point, but different from 0,
finally, from Fig. \ref{subfig:roc-worst-NN}, the ROC curve and AUC confirms the
poor performance of this classifier. Finally, the Sensitivity-Specificity
results (see Table \ref{tab:L123}) show that this NN has adequate values for the
specificity but deplorable values for the sensitivity.

@@

As for future work, the hypothesis and results here presented could be reviewed
by an expert, since this work is preliminary and the authors are no experts in
Mental Health. Additionally, a better review of sophisticated techniques for
visualizing learning machines could be done, since the 2-dimensional embedding
of the decision function (contours) was more of an heurisitc approach proposed
by the authors and it could be somehow improved. Finally, the inclusion of more
metrics for evaluating the performance of the classifiers can be included as
well, since only the Sensitivity-Specificity was done.

>>>>

Given that the sample size is fixed, the analysis for the ``Probably
Approximately Correct'' (PAC) learning was not straight-forward, as the
classifier's maximum permitted error $\epsilon$ was estimated from the classic
formulae, rather than calculating the adequate sample size for a given $\delta$.

The PAC learning here applied is based on the ideas presented in
\parencite{kubat2017}. Let $N$ be number of samples in the training set,
$\epsilon$ is the classifier maximum permitted error rate, $\delta$ is the
probability that a classifier with error rate greater than $\epsilon$ is
\text{error-free} on the training set, $m$ be the dimension of each training
pattern, and let $VC$ be the Vapnik-Chervonenkis dimension of the given
classifier. Hence, for learning in continuous domains, the lower bound for $N$
is

@@

Furthermore, it is important to notice that the dataset is 6-dimensional, which
clearly brings problems for visualizing the learning machines' outputs and
decision boundaries. In this paper, the decision functions of the fitted
learning machines will be presented using contour plots of the projected data.
This process will better described: each 6-dimensional pattern generates an
output whenever it is evaluated in the fitted model. The idea is to generate an
equally-spaced grid in $[-1,1]^6$ and evaluate each point on this model, in
order to plot the contour of the shattering function. The main issue is that a
direct projection into a 2-dimensional space of each point on the grid is not
necessarily a function, since several points can have the same projection, but
different values of the shattering function. In order to address this issue, the
mean of the points that mapped into the same projection is taken, giving an
proxy of the general behavior in high dimensions. This ``mean'' projection
contour plots also include the projection of the training data points into the
same coordinates and scattered with the real labels, giving an idea of the
classification performance in high dimensions. Remark: as stated, the survey
data also includes the outputs of the Fuzzy Inference System; thus, the contour
plots are presented only for relating first input with the first input, second
with the second output and so on. This was done aiming at two objectives:
keeping a reasonable number of plots in the document, and giving a more
intuitive interpretation to these plots.

@@

The same idea from the SVM classifiers for plotting the contour of the decision
boundaries of the fitted model was used, plotting the training dataset with the
real labels and the contour for the decision function.

Additionally, it is well-known that the bound for the training dataset size is
given by
%
\begin{equation}
  N\geq \dfrac{\ln(2)}{2\epsilon^2}\left[(2^k-1)(1+\log_2 m) + 1 + \ln\left(
      \dfrac{1}{\delta}\right)\right],
\end{equation}
%
but again, the size of the dataset is fixed. Thus, $\epsilon$ was calculated by
fixing the tree depth $k=4$ and setting $\delta=0.3$ yielding a value of
$\epsilon=0.5381$, which is a suitable value for the classifier error.

@@

The results are presented using three learning curves: the sum of local
gradients per layer at the end of each epoch, the average error energy and a
Receiver Operating Characteristic (ROC) curve, comparing the obtained results
with the real labels.
%
\begin{itemize}
  \item The sum of gradients per layer at the end of each epoch is an approach
        visualizing if the training algorithm is solving the optimization
        problem, since the sum of local gradients must converge to zero as each
        gradient converges to zero. It is well-known that the vanish of
        gradients is a necessary condition for optimality and the training
        algorithm is just an optimization problem minimizing the output error
        (cost function).
  \item The average error energy is the cost function of the optimization
        problem. Let $y_i\in\mathbb{R}^n$ be the real output of a given input
        pattern $x_i\in\mathbb{R}^m$ and let $\hat{y}_i\in\mathbb{R}^n$ be the
        output of the NN for the same input pattern. Denoting each component by
        superscripts, the average error energy $\xi_{av}$ on each epoch is then
        computed as
        %
        \begin{equation}
          \xi_{av}=\dfrac{1}{2}\sum_{i=1}^{N}\sum_{k=1}^n\left(y^k_i-\hat{y}^k_i\right)^2,
        \end{equation}
        where $N$ is the training sample size. Thus, it is clear that
        $\xi_{av}\rightarrow0$ is a good initial approach for evaluating the
        learning performance.
  \item The ROC curve is a tool that helps visualize the correct classifications
        of the learning machine by plotting the true positive rate on the
        vertical axis, and false positive rate on the horizontal axis. The Area
        Under the Curve (AUC) is an indicator of how good the data is being
        classified by the model, with higher values being usually better.
\end{itemize}

\subsection{Sensitivity-Specificity}
The sensitivity is a ``measure of the proportion of actual positive cases that
got predicted as positive (true positive)'' and the specificity is ``measure of
proportion of actual negatives that got predicted as negatives (true negatives)
'' \parencite{sens}. This values are, hence, calculated using the formulae
%
\begin{align}
  \mathrm{Sensitivity}=&\dfrac{TP}{TP+FN}, \\
  \mathrm{Specificity} =& \dfrac{TN}{TN+FP},
\end{align}
%
where TP stands for True Positives, FN for False Negatives, TN for True
Negatives and FP for False Negatives, which are extracted from a confusion
matrix.

\section{Results}\label{sec:res}
For giving an idea of the dataset treated in this work (visualization), the data
was embedded into a 2-dimensional space using the T-student Stochastic
Neighborhood Embedding (TSNE) \href{https://github.com/lejon/TSne.jl}{in Julia}
\parencite{maaten2008}. In Fig. \ref{fig:emb_dat}, the embedded dataset is
presented.

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/embedded-data.pdf}
  \caption{Embedded dataset.}
  \label{fig:emb_dat}
\end{figure}

@@

\subsection{Neural Networks}
Recall that only the worst and 2 best NN are kept for presenting the results.
The worst NN (in the sense of the objective function) was trained with
$\eta=0.9$ and it has a $\mathcal{L}=[1,2,3]$ architecture in the hidden layers.
The learning curves for this MLP are presented in Fig. \ref{fig:NN-worst}. This
architecture is expected to perform poorly, at least for the given number of
epochs, since it is attempting to summarize all 6 inputs in a one dimensional
value for the first hidden layer and then casting this value into higher
dimensions. It is unknown for the authors if this machine can learn the data at
all, since the behavior of the backpropagation algorithm is unpredictable as
epochs advance. Note from Fig. \ref{subfig:grad-worst-NN}, the gradients are
stable in the epoch horizon defined, from Fig. \ref{subfig:error-worst-NN} that
the average energy error seems to reach a stable point, but different from 0,
finally, from Fig. \ref{subfig:roc-worst-NN}, the ROC curve and AUC confirms the
poor performance of this classifier. Finally, the Sensitivity-Specificity
results (see Table \ref{tab:L123}) show that this NN has adequate values for the
specificity but deplorable values for the sensitivity.

@@

As for future work, the hypothesis and results here presented could be reviewed
by an expert, since this work is preliminary and the authors are no experts in
Mental Health. Additionally, a better review of sophisticated techniques for
visualizing learning machines could be done, since the 2-dimensional embedding
of the decision function (contours) was more of an heuristic approach proposed
by the authors and it could be somehow improved. Finally, the inclusion of more
metrics for evaluating the performance of the classifiers can be included as
well, since only the Sensitivity-Specificity was done.
